Evolutionary rescue and reintroduction of resistant frogs allows recovery in the presence of a lethal fungal disease

### 2023-05-19 doc/manuscript/translocation_pnas.tex
Update PNAS version and add biorxiv version

\begin{abstract}
Escalating anthropogenic change is devastating global biodiversity. To
avoid extinction, affected populations will often need to adapt to the
altered environment. The process of rapid evolutionary change that
increases the frequency of adaptive alleles and thereby restores
positive population growth is termed ``evolutionary rescue''. In
addition to allowing recovery of extant populations, individuals from
rescued populations could also be used to reestablish extirpated
populations. We evaluated this scenario in the once-common mountain
yellow-legged (MYL) frog, which is threatened with extinction by the
human-facilitated emergence of a lethal fungal pathogen
(\emph{Batrachochytrium dendrobatidis}; ``Bd''). Although most MYL frog
populations are extirpated following disease outbreaks, some persist and
eventually begin to recover. We show that MYL frogs have undergone
substantial evolutionary change following disease outbreaks, and that
these changes are associated with increased resistance/tolerance to Bd
infection in naturally recovering populations. Large-scale
reintroduction of frogs from these populations resulted in the
establishment of reproducing populations at most recipient sites despite
ongoing disease. In addition, results from viability modeling suggest
that established populations have a low probability of extinction over
50 years. Collectively, these results provide one of the first examples
of how evolutionary rescue and the reintroduction of resistant/tolerant
individuals can allow the landscape-scale recovery of disease-impacted
species. This example has potentially broad implications for the many
taxa worldwide that are threatened with extinction by anthropogenic
stressors.
\end{abstract}

\dates{This manuscript was compiled on \today}
\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

@@

\dropcap{H}uman activities are intensifying threats to global biodiversity
\citep{ceballos2015}, with important implications for ecosystem
resilience and human welfare \citep{naeem2009}. To persist, affected
taxa will often need to adapt to their altered environment or face
decline and extinction. ``Evolutionary rescue'' describes rapid
evolutionary change that increases the frequency of adaptive alleles and
thereby restores positive population growth \citep{carlson2014}. In
addition to preventing extinction of declining populations, this process
could also result in the seeding of nearby empty habitats and allow
population reestablishment and recovery. Although this overall scenario
is potentially of critical importance for the conservation of an
ever-growing number of threatened taxa, descriptions of its operation in
the wild are lacking.

One consequence of human alteration of the biosphere is an increase in
emerging infectious diseases \citep{jones2008, fisher2012}. Such
diseases pose a severe threat to wildlife populations
\citep{daszak2000}, and have caused dramatic declines and extinctions in
a wide range of taxa
\citep{hewson2014, samuel2015, scheele2019, cunningham2021}. Amid this
global loss of biodiverity, some populations decimated by disease have
subsequently recovered to varying extents
\citep{newell2013, voyles2018, knapp2016}. Although we know relatively
little about the mechanisms underlying recovery
\citep{brannelly2021, russell2020}, host evolution is one possibility
\citep{carlson2014, searle2020}. This process is known from a diversity
of wildlife populations
\citep{savage2016, epstein2016, gignoux-wolfsohn2021, holland2022}, and
often results in increased host resistance and/or tolerance (see
\citep{medzhitov2012} for definitions). However, linking these
evolutionary responses to population recoveries is difficult and few
clear examples exist. In addition to promoting recovery of
disease-impacted populations, the resistant/tolerant individuals in such
populations could potentially be used to reestablish populations that
were extirpated following disease emergence
\citep{mendelson2019, brannelly2021}. To date, the use of
resistant/tolerant individuals for species recovery is rare \citep[but
see][]{joseph2018}, and studies that evaluate its potential utility are
critically needed.

@@

We collected DNA samples via buccal swabbing \citep{broquet2007} from 53
\emph{Rana muscosa}/\emph{Rana sierrae} individuals: 24 from four
Bd-naive populations, and 29 from five recovering populations. These
populations were located in the southern Sierra Nevada, from northern
Yosemite National Park to northern Sequoia National Park
(Figure~\ref{fig-allelefrequencies}). Samples were collected from 5-6
frogs per population. To minimize potential confounding effects caused
by variation in frog genotypes across latitude
\citep[e.g.,][]{rothstein2020}, we selected sampling sites such that
both population types were represented across similar latitudinal
ranges. DNA was extracted following Qiagen DNEasy manufacturer's
protocols. We sequenced the samples using an exome capture assay as
described in Byrne et al.~(in review). Briefly, genomic libraries were
prepared and captured using a custom Nimblegen capture pool. Capture
baits were designed based on the coding regions of the \emph{R. muscosa}
transcriptome (GenBank accession GKCT00000000). Captured libraries were
then pooled and sequenced on a NovaSeq 6000 150PE Flow Cell S1 at the
Vincent J. Coates Genomics Sequencing Lab at UC Berkeley.

@@

Raw reads were filtered for adapters and contaminants using fastp
\citep{chen2018} and aligned to the \emph{Rana muscosa} genome
\citep[NCBI SRA: SRS6533475, reference =][]{hon2020} with repetitive
elements masked using bwa (``mem'' mode) \citep{li2013}. Exact PCR
Duplicates were marked using Picard. Variants were then called following
GATK best practices (v.4.2.0.0, \citep{van2020}). Briefly, raw variants
were called for each sample using HaplotypeCaller and combined using
CombineGVCFs. Next, genotypes were jointly called using GenotypeGVCFs.
Variants were then hard filtered using gatk VariantFiltration using the
following parameters to remove low-quality sites: QD \textless{} 2.0, FS
\textgreater{} 40.0, SOR \textgreater{} 3.0, MQ \textless{} 50.0,
MQRankSum \textgreater{} 3.0, MQRankSum \textless{} -3.0, ReadPosRankSum
\textgreater{} 3.0, ReadPosRankSum \textless{} -3.0. This initial filter
resulted in 1,595,206 variant sites across 53 individuals.

We then further filtered our dataset at the individual and variant
level. First, we trimmed our variants to only include those with minor
allele frequency \textgreater{} 0.03, a maximum depth of 250 and minimum
depth of 5, a minimum genotype quality of 20, and a maximum missing
proportion of 0.5. This filter resulted in 427,038 sites, of which
353,172 were single nucleotide polymorphisms (SNPs) and 73,866 were
insertions or deletions (INDELS). Finally, we trimmed samples with an
average depth across filtered sites \textless{} 7x. Our final dataset
included 50 samples (23 from naive and 27 from recovering populations)
with an average depth of 16.7x (range = 7.4x -- 26.1x).

@@

To visualize the genomic relationship of our populations we conducted a
principal components analysis (PCA) using the glPCA function in the
adegenet R package \citep{jombart2008}. To detect regions of the genome
under selection between naive and recovering populations, we used two
approaches: (1) a multivariate linear mixed model to evaluate individual
variants (SNPs and INDELs), and (2) a splined window analysis to
evaluate larger genomic regions. For the variant analysis, we first used
a stringent data filter to include only variants with \textless{} 5\%
missing data (missing for no more than 2 individuals), and then
calculated the likelihood ratio statistic for the resulting set of
148,307 high quality variants across 127 contigs using GEMMA. GEMMA
calculates and incorporates a relatedness matrix for input samples,
allowing us to account for relatedness and population structure when
calculating likelihood ratio statistics. We identified variants showing
different allele frequencies between naive versus recovering populations
(``outliers'') using a Bonferroni- corrected significance level of 0.01.
We visualized the results using a Manhattan plot and qqplot. We
developed a more liberal set of outlier variants using a Bonferroni-
corrected significance level of 0.05 and used this set solely for the
gene ontology (GO) analysis (see below and SI).

For each outlier variant, we determined whether the variant was
synonymous (protein sequence the same for each variant) or
non-synonymous (protein sequence differs between variants), and where in
the gene it was located. To do this, we first extracted the reference
genome sequence surrounding the variant using the bedtools ``getfasta''
function \citep{quinlan2010}. Next, we re-annotated each sequence using
blast to get the predicted gene location based on the closest annotated
reference \citep{altschul1997}. We then translated each variant to amino
acids and aligned this translation to that of the gene annotation to
ensure proper frame of reference using Geneious \citep{kearse2012}.
After ensuring proper translation, we characterized variants as within
or outside the coding sequence of the gene and as either synonymous or
non-synonymous.

In the splined window analysis, we identified outlier regions using
\emph{F\textsubscript{ST}} and differences in nucleotide diversity
(\(\pi_{diff}\)) between naive and recovering populations. First, we
calculated per-site \emph{F\textsubscript{ST}} between the naive and
recovering individuals for all bi-allelic SNPs in the 30 largest contigs
(98\% of all SNPs) using VCFtools \citep{danecek2011}. Next, we
calculated per-site nucleotide diversity \(\pi\) separately for
individuals from the naive and recovering populations using VCFtools,
then calculated \(\pi_{diff}\) for each population
(\(\pi_{diff} = \pi_{naive} - \pi_{recovering}\)). We concatenated the
values for \emph{F\textsubscript{ST}} and \(\pi_{diff}\) in order of
size-sorted chromosome number and adjusted the SNP position based on the
relative position in the genome. We then used the WINGEN R package
\citep{bishop2023} to conduct a splined discrete window analysis for
\emph{F\textsubscript{ST}} and \(\pi_{diff}\). This method calculates
where non-overlapping window boundaries should occur by identifying
inflection points in the fitted spline, therefore balancing false
positive and false negative results that occur using other window-based
methods \citep{beissinger2015} . This method also calculates a
W-statistic allowing for outlier identification. We identified outliers
as those with a W-statistic greater than 4 standard deviations above the
mean for F\textsubscript{ST} or above/below the mean for \(\pi_{diff}\).
Outlier regions were then identified as those that were outliers in both
analyses, meaning that they showed 1) high differentiation between naive
and recovering populations, and 2) differential patterns of nucleotide
diversity in the same region. Finally, we extracted gene transcripts
mapped within each region and retrieved annotation for that region using
blast.

Using the liberal set of variant outliers, we performed a statistical
overrepresentation test (Fisher's exact test) using Blast2GO
\citep{conesa2005}. First, we retrieved the Blast hits and mapped GO
terms for each gene in our targeted transcriptome. Next, we compared the
29 unique outlier genes from the liberal set of outlier variants to the
complete set of genes in our target transcriptome to determine if any GO
biological functions, molecular functions, or cellular processes were
overrepresented. We then repeated this process for the set of 33 genes
located in the overlapping \emph{F\textsubscript{ST}} and \(\pi_{diff}\)
splined windows. Code used for genomic analyses and to figures is
available from the following GitHub repository:
\url{https://github.com/allie128/mylf-selection}.

@@

For the 24 translocations we conducted, we identified donor populations
from which adult frogs (\(\geq\) 40 mm snout-vent length) could be
collected using several years of visual encounter surveys (VES) and skin
swab collections \citep{knapp2016}, as well as landscape genetic
information \citep{poorten2017}. The populations that we selected
contained several hundred \emph{R. sierrae} adults and hundreds or
thousands of tadpoles. These relatively high abundances were the result
of recent increases following previous Bd-caused declines
\citep{knapp2016}. As is typical for recovering MYL frog populations, Bd
prevalence in the donor populations was high (0.69--0.96) and Bd load
(median log\textsubscript{10}(load) = 3.06--3.78 ITS copies) was two or
more orders of magnitude below the level at which frog mortality is
expected (log\textsubscript{10}(load) \(\approx\) 5.78
copies)\citep{vredenburg2010, joseph2018}. Recipient sites to which
frogs were translocated were chosen based on previous \emph{R. sierrae}
presence (determined from VES and/or museum records) or characteristics
that suggested high quality habitat for this species \citep{knapp2005}.

We conducted 1--4 translocations per site (Figure S3)
and each translocated cohort included 18 to 99 frogs (median = 30).
During each translocation, adult frogs were collected from the donor
population and measured, weighed, swabbed, and PIT tagged. Frogs were
transported to the recipient site either on foot or via helicopter.
Following release, translocated populations were visited approximately
once per month during the summer active season (median number of visits
per summer = 2; range = 1--10) and assessed using diurnal CMR surveys
and VES. CMR surveys allowed quantification of adult survival, adult
population size, and recruitment of new adults, and VES provided
estimates of tadpole and juvenile abundance. During 2006-2012, CMR
surveys were conducted on a single day (primary period) per site visit,
during which all habitats were searched repeatedly for adult frogs.
Frogs were captured using handheld nets, identified via their PIT tag
(or tagged if they were untagged), measured, weighed, swabbed, and
released at the capture location. During 2013-2022, we generally used a
robust design in which all habitats were searched during several
consecutive days (median number of secondary periods per primary period
= 3; range = 3--7), and frogs were processed as described above.
However, when the number of frogs detected on the first survey day was
zero or near zero, we typically conducted only a single-day CMR survey.
When using a robust design, within a primary period, frogs that were
captured during more than one secondary period were measured, weighed,
and swabbed during the first capture, and during subsequent captures
were only identified and released. Frog skin swabs were analyzed using
standard Bd DNA extraction and qPCR methods. For details of survey and
swab analysis methods, see \citep{joseph2018}.

During each site visit, we conducted VES either immediately before CMR
surveys or during the first day of CMR surveys. VES was conducted by
walking the entire water body perimeter, first 100 m of each inlet and
outlet stream, and any fringing ponds and wetlands, and counting all
\emph{R. sierrae} tadpoles and juveniles. These \emph{R. sierrae} life
stages have high detectablility, and counts are highly repeatable and
provide estimates of relative abundance \citep{knapp2000}.

@@

To identify important predictors of frog survival following
translocation, we developed linear and multilevel models in a Bayesian
framework. We performed all analyses with the rstanarm package
\citep{rstanarm2022} and R Statistical Software
\citep[v4.4.4,][]{rsoftware2022}, and used default, weakly informative
priors. Starting models included all relevant population-level effects
and, when data contained possible hierarchical structure, we included
group-level effects in the full model. Within each analysis, we
evaluated model fit using visualizations of leave-one-out (``LOO'')
probability integral transformations \citep{gelman2013, gabry2019}, and
compared fits of models with and without group-level effects, and with
differing group-level effects using LOO cross-validation
\citep{vehtari2016}. Both approaches utilized the loo package
\citep{vehtari2022}.

We conducted two analyses. In the first, we addressed the more general
question of whether survival of frogs in the first translocation to a
site predicted the survival of frogs in subsequent translocations.
Group-level effects included site\_id or translocation\_id. The second
analysis focused specifically on identifying predictors of
post-translocation frog survival using a wide range of variables
describing characteristics of frogs, sites, and translocations
(predictor variables: Bd load (bd\_load), winter severity in the year
following translocation (snow\_t1), site elevation (elevation), donor
population (donor\_72996), sex (sex\_male), frog size (length), winter
severity in the year of translocation (snow\_t), day of year on which a
translocation was conducted (day), and translocation order
(order\_first)). In both analyses we used 1-year survival as a single
metric of survival that was available for all translocated populations
regardless of how many years of post-translocation data were available
(range = 3-\/-16). Group-level effects included site\_id,
translocation\_id, or translocation\_id nested within site\_id. In both
analyses, the response variable, probability of 1-year frog survival,
was rounded to integer values to produce a binomial outcome, and all
models used a bernoulli family. For all models, we used default priors,
four chains, and 5000 iterations each for warmup and post-warmup. We
checked convergence of the MCMC algorithm using trace plots and Rhat.
Code to create the datasets and fit models is available from the
following GitHub repository:
\url{https://github.com/SNARL1/translocation}.

@@

We analyzed skin swabs using standard Bd DNA extraction and qPCR methods
\citep[see SI for details]{boyle2004}. To assess the magnitude of
changes in Bd load on frogs following translocation, we compared Bd
loads measured before versus after translocation. Before-translocation
loads were quantified using skin swabs collected from all
to-be-translocated frogs at the donor site on the day before or the day
of the translocation. After-translocation Bd loads were based on all
swabs collected from translocated frogs at the recipient site in the
year of and the year following translocation. Individual frogs and their
associated Bd loads were included in the dataset only if frogs were
captured at the recipient site at least once during the year 0-1 period.
We used an analysis workflow similar to that described above. Briefly,
the response variable was before-translocation load and the predictor
variable was after-translocation load. We compared models with and
without group-level effects, and group-level effects included
pit\_tag\_id, pit\_tag\_id nested within site\_id, or pit\_tag\_id
nested within translocation\_id. All models used a negative binomial
family. Priors, chains, iterations, and checks were as described above.
Code to create the dataset and fit models is available from the
following GitHub repository:
\url{https://github.com/SNARL1/translocation}.

@@

In this model we ignore density-dependent recruitment because we were
interested in the growth of the population from an initial
reintroduction and whether this growth was sufficient to prevent
extinction over 50 years following the introduction. We also did not
directly consider the dynamics of Bd in this model. We made this
decision because (i) translocated populations are infected with Bd at
high prevalence \citep{joseph2018}, and (ii) host density does not seem
to play a significant role in multi-year Bd infection dynamics in this
system \citep{wilber2022}. Thus, ignoring Bd infection dynamics and
instead assuming all host vital rates are in the presence of high Bd
prevalence significantly simplifies the model without much loss of
realism.

@@

\caption{\label{fig-allelefrequencies}Evidence for selection in
recovering MYL frog populations at the landscape scale. For each of the
9 naive and recovering MYL frog populations, pie charts show allele
frequencies for the 11 outlier SNPs from 7 distinct genes: (A)
LOC108802036, (B) TCF19, VARS, MAP1S, TYRP1, SRFBP1, and (C) RIN3.
Charts are superimposed on a map of the Sierra Nevada study area, with
Yosemite, Kings Canyon, and Sequoia National Parks (from north to south)
shown in dark gray, and the range boundary for MYL frogs in this portion
of the Sierra Nevada shown in light gray. The inset map locates the
study area in California.}

@@

{\centering \includegraphics[width=\textwidth]{figures/translocation_survival_bysiteid.png}

>>>>

\begin{abstract}
Vast alteration of the biosphere by humans is causing a sixth mass
extinction. Adaptation to modified environments often is the only means
for species to persist, but the extent to which such ``evolutionary
rescue'' can prevent or reverse biodiversity loss is largely unknown.
Using results from genomic analyses, a 15-year reintroduction effort,
and population modeling, we provide a compelling example of evolution
reversing the declines of an imperiled amphibian and allowing the
reestablishment of extirpated populations. The once-common mountain
yellow-legged (MYL) frog is threatened with extinction by the
human-facilitated emergence of a lethal fungal pathogen
(\emph{Batrachochytrium dendrobatidis}; ``Bd''). Although most MYL frog
populations are extirpated following disease outbreaks, some persist and
eventually begin to recover. Using an exome capture approach, we show
that MYL frogs have undergone substantial evolutionary change following
disease outbreaks. Moreover, in naturally recovering populations
observed changes in immune function loci are associated with increased
resistance/tolerance to Bd infection. Large-scale reintroduction of
frogs from rescued populations resulted in the establishment of
reproducing populations despite ongoing disease. In addition, results
from viability modeling suggest that established populations have a low
probability of extinction over 50 years. Collectively, these results
provide a rare example of how evolutionary rescue and the reintroduction
of resistant/tolerant individuals can allow the landscape-scale recovery
of disease-impacted species. This example has potentially broad
implications for the many taxa worldwide that are threatened with
extinction by anthropogenic stressors.
\end{abstract}

\dates{This manuscript was compiled on \today}
% \doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}

@@

\dropcap{H}uman activities are intensifying threats to global biodiversity
\citep{ceballos2015}, with important implications for ecosystem
resilience and human welfare \citep{naeem2009}. To persist, affected
taxa will often need to adapt to their altered environment or face
decline and possible extinction. The process whereby rapid evolutionary
change increases the frequency of adaptive alleles and restores positive
population growth is known as ``evolutionary rescue''
\citep{carlson2014}. In addition to preventing extinction of declining
populations, individuals from rescued populations could also disperse,
or be reintroduced, to empty habitats and promote larger-scale recovery.
Although this overall scenario is potentially of critical importance for
the conservation of an ever-growing number of imperiled taxa,
descriptions of its occurrence and effectiveness in the wild are
lacking.

One consequence of human alteration of the biosphere is an increase in
emerging infectious diseases \citep{jones2008, fisher2012}. Such
diseases pose a severe threat to wildlife populations
\citep{daszak2000}, and have caused dramatic declines and extinctions in
a wide range of taxa, including echinoderms, mammals, birds, and
amphibians \citep{hewson2014, samuel2015, scheele2019, cunningham2021}.
Amid this global loss of biodiversity, some populations decimated by
disease have subsequently recovered to varying extents
\citep{newell2013, voyles2018, knapp2016}. Although we know relatively
little about the mechanisms underlying recovery
\citep{brannelly2021, russell2020}, host evolution is one possibility
\citep{carlson2014, searle2020}. This process is known from a diversity
of wildlife populations
\citep{savage2016, epstein2016, gignoux-wolfsohn2021, holland2022}, and
often results in increased host resistance and/or tolerance (see
\citep{medzhitov2012} for definitions). However, linking these
evolutionary responses to population recoveries is difficult and few
clear examples exist. In addition to promoting recovery of
disease-impacted populations, the resistant/tolerant individuals in such
populations could potentially be used to reestablish populations that
were extirpated following disease emergence
\citep{mendelson2019, brannelly2021}. To date, the use of
resistant/tolerant individuals for species recovery is rare \citep[but
see][]{joseph2018}, and studies that evaluate its potential utility are
critically needed.

@@

We collected DNA samples via buccal swabbing \citep{broquet2007} from 53
\emph{Rana muscosa}/\emph{Rana sierrae} individuals: 24 from four naive
populations, and 29 from five recovering populations. These populations
were located in the southern Sierra Nevada, from northern Yosemite
National Park to northern Sequoia National Park
(Figure~\ref{fig-allelefrequencies}). Samples were collected from 5-6
frogs per population. To minimize potential confounding effects caused
by variation in frog genotypes across latitude
\citep[e.g.,][]{rothstein2020}, we selected sampling sites such that
both population types were represented across similar latitudinal
ranges. DNA was extracted following Qiagen DNEasy manufacturer's
protocols. We sequenced the samples using an exome capture assay as
described in Byrne et al.~(in review). Briefly, genomic libraries were
prepared and captured using a custom Nimblegen capture pool. Capture
baits were designed based on the coding regions of the \emph{R. muscosa}
transcriptome (GenBank accession GKCT00000000). Captured libraries were
then pooled and sequenced on a NovaSeq 6000 150PE Flow Cell S1 at the
Vincent J. Coates Genomics Sequencing Lab at UC Berkeley.

@@

Raw reads were filtered for adapters and contaminants using fastp
\citep{chen2018} and aligned to the \emph{Rana muscosa} genome (NCBI
SRA: SRS6533475 \citep{hon2020}), with repetitive elements masked using
bwa (``mem'' mode) \citep{li2013}. Exact PCR duplicates were marked
using Picard. Variants were then called following GATK best practices
(v.4.2.0.0 \citep{van2020}). Briefly, raw variants were called for each
sample using HaplotypeCaller and combined using CombineGVCFs. Next,
genotypes were jointly called using GenotypeGVCFs. Variants were then
hard filtered using gatk VariantFiltration using the following
parameters to remove low-quality sites: QD \textless{} 2.0, FS
\textgreater{} 40.0, SOR \textgreater{} 3.0, MQ \textless{} 50.0,
MQRankSum \textgreater{} 3.0, MQRankSum \textless{} -3.0, ReadPosRankSum
\textgreater{} 3.0, ReadPosRankSum \textless{} -3.0. This initial filter
resulted in 1,595,206 variant sites across 53 individuals.

We then further filtered our dataset at the individual and variant
level. First, we trimmed our variants to only include those with minor
allele frequency \textgreater{} 0.03, a maximum depth of 250 and minimum
depth of 5, a minimum genotype quality of 20, and a maximum missing
proportion of 0.5. This filter resulted in 427,038 sites, of which
353,172 were SNPs and 73,866 were INDELS. Finally, we trimmed samples
with an average depth across filtered sites \textless{} 7x. Our final
dataset included 50 samples, 23 from naive and 27 from recovering
populations, with an average depth of 16.7x (range = 7.4x -- 26.1x).

@@

To visualize the genomic relationship of our populations we conducted a
PCA using the glPCA function in the adegenet R package
\citep{jombart2008}. To detect regions of the genome under selection
between naive and recovering populations, we used two approaches: (1) a
multivariate linear mixed model to evaluate individual variants (SNPs
and INDELs), and (2) a splined window analysis to evaluate larger
genomic regions. For the variant analysis, we first used a stringent
data filter to include only variants with \textless{} 5\% missing data
(missing for no more than 2 individuals), and then calculated the
likelihood ratio statistic for the resulting set of 148,307 high quality
variants across 127 contigs using GEMMA \citep{zhou2014}. GEMMA
calculates and incorporates a relatedness matrix for input samples,
allowing us to account for relatedness and population structure when
calculating likelihood ratio statistics. We identified variants showing
different allele frequencies between naive versus recovering populations
(``outliers'') using a Bonferroni- corrected significance level of 0.01.
We visualized the results using a Manhattan plot and qqplot. We
developed a more liberal set of outlier variants using a Bonferroni-
corrected significance level of 0.05 and used this set solely for the
gene ontology (GO) analysis (see below and SI Appendix: Dataset S1, S2).

In the analysis of individual variants, for each outlier variant we
determined whether the variant was synonymous (protein sequence the same
for each variant) or non-synonymous (protein sequence differs between
variants), and where in the gene it was located. To do this, we first
extracted the reference genome sequence surrounding the variant using
the bedtools ``getfasta'' function \citep{quinlan2010}. Next, we
re-annotated each sequence using BLAST to get the predicted gene
location based on the closest annotated reference \citep{altschul1997}.
We then translated each variant to amino acids and aligned this
translation to that of the gene annotation to ensure proper frame of
reference using Geneious \citep{kearse2012}. After ensuring proper
translation, we characterized variants as within or outside the coding
sequence of the gene and as either synonymous or non-synonymous.

In the splined window analysis, we identified outlier regions using
\emph{F\textsubscript{ST}} and differences in nucleotide diversity
(\(\pi_{diff}\)) between naive and recovering populations. First, we
calculated per-site \emph{F\textsubscript{ST}} between the naive and
recovering individuals for all bi-allelic SNPs in the 30 largest contigs
(98\% of all SNPs) using VCFtools \citep{danecek2011}. Next, we
calculated per-site nucleotide diversity \(\pi\) separately for
individuals from the naive and recovering populations using VCFtools,
then calculated \(\pi_{diff}\) for each population
(\(\pi_{diff} = \pi_{naive} - \pi_{recovering}\)). We concatenated the
values for \emph{F\textsubscript{ST}} and \(\pi_{diff}\) in order of
size-sorted chromosome number and adjusted the SNP position based on the
relative position in the genome. We then used the GenWin R package
\citep{beissinger2015} to conduct a splined discrete window analysis for
\emph{F\textsubscript{ST}} and \(\pi_{diff}\). This method calculates
where non-overlapping window boundaries should occur by identifying
inflection points in the fitted spline, therefore balancing false
positive and false negative results that occur using other window-based
methods \citep{beissinger2015}. This method also calculates a
W-statistic allowing for outlier identification. We identified outliers
as those with a W-statistic greater than 4 standard deviations above the
mean for \emph{F\textsubscript{ST}} or above/below the mean for
\(\pi_{diff}\). Shared outliers were then identified as those that were
outliers in both analyses, meaning that they showed (1) high
differentiation between naive and recovering populations, and (2)
differential patterns of nucleotide diversity in the same region.
Finally, we extracted gene transcripts mapped within each region and
retrieved annotation for that region using BLAST (SI Appendix: Datasets
S3, S4).

Using the liberal set of variant outliers, we performed a statistical
overrepresentation test (Fisher's exact test) using Blast2GO
\citep{conesa2005}. First, we retrieved the BLAST hits and mapped GO
terms for each gene in our targeted transcriptome. Next, we compared the
30 unique outlier genes from the liberal set of outlier variants to the
complete set of genes in our target transcriptome to determine if any GO
biological functions, molecular functions, or cellular processes were
overrepresented. We then repeated this process for the set of 33 genes
located in the overlapping \emph{F\textsubscript{ST}} and \(\pi_{diff}\)
splined windows. Code used for genomic analyses and to create figures is
available from the following GitHub repository:
\url{https://github.com/allie128/mylf-selection}.

@@

For the 24 translocations we conducted, we identified donor populations
from which adult frogs (\(\geq\) 40 mm snout-vent length) could be
collected using several years of visual encounter surveys (VES) and skin
swab collections \citep{knapp2016}, and results from population genetic
analyses \citep{poorten2017}. The populations that we selected contained
several hundred \emph{R. sierrae} adults and hundreds or thousands of
tadpoles. These relatively high abundances were the result of recent
increases following previous Bd-caused declines \citep{knapp2016}. As is
typical for recovering MYL frog populations, Bd prevalence in the donor
populations was high (0.69--0.96) and Bd load (median
log\textsubscript{10}(load) = 3.06--3.78 ITS copies) was two or more
orders of magnitude below the level at which frog mortality is expected
(log\textsubscript{10}(load) \(\approx\) 5.78
copies)\citep{vredenburg2010, joseph2018}. Recipient sites to which
frogs were translocated were chosen based on previous \emph{R. sierrae}
presence (determined from VES and/or museum records) or characteristics
that suggested high quality habitat for this species \citep{knapp2005}.

We conducted 1--4 translocations per site (Figure S3) and
each translocated cohort included 18 to 99 frogs (median = 30). In
preparation for each translocation, adult frogs were collected from the
donor population and measured, weighed, swabbed, and PIT tagged. Frogs
were transported to the recipient site either on foot or via helicopter.
Following release, translocated populations were visited approximately
once per month during the summer active season (median number of visits
per summer = 2; range = 1--10) and assessed using diurnal CMR surveys
and VES. CMR surveys allowed estimation of adult survival, adult
population size, and recruitment of new adults, and VES provided
estimates of tadpole and juvenile abundance. During 2006-2012, CMR
surveys were conducted on a single day (primary period) per site visit,
during which all habitats were searched repeatedly for adult frogs.
Frogs were captured using handheld nets, identified via their PIT tag
(or tagged if they were untagged), measured, weighed, swabbed, and
released at the capture location. During 2013-2022, we generally used a
robust design in which all habitats were searched during several
consecutive days (median number of secondary periods per primary period
= 3; range = 3--7), and frogs were processed as described above.
However, when the number of frogs detected on the first survey day was
zero or near zero, we typically conducted only a single-day CMR survey.
When using a robust design, within a primary period, frogs that were
captured during more than one secondary period were measured, weighed,
and swabbed during the first capture, and during subsequent captures
were only identified and released. Frog skin swabs were analyzed using
standard Bd DNA extraction and qPCR methods. For details of survey and
swab analysis methods, see \citep{joseph2018}.

During each site visit, we conducted VES either immediately before CMR
surveys or during the first day of CMR surveys. VES was conducted by
walking the entire water body perimeter, first 100 m of each inlet and
outlet stream, and any fringing ponds and wetlands, and counting all
\emph{R. sierrae} tadpoles and juveniles. These \emph{R. sierrae} life
stages have high detectability, and counts are highly repeatable and
provide estimates of relative abundance \citep{knapp2000}.

\hypertarget{frog-counts-and-reproductive-success}{%
\subsection*{Frog counts and reproductive
success}\label{frog-counts-and-reproductive-success}}

For each of the translocated populations, we used the presence of
tadpoles and/or juveniles from VES and captures of new recruits (i.e,
untagged adults) in CMR surveys to provide two measures of successful
reproduction. To calculate the proportion of years in which
tadpoles/juveniles were present at a site, we excluded surveys conducted
in the year of the initial translocation to that site. This exclusion
accounted for the fact that all translocations were conducted after the
breeding period and reproduction would therefore not have occurred until
the following year. Similarly, to calculcate the proportion of years in
which new recruits were present at a site, we excluded surveys conducted
during the 3 years following the initial translocation. This accounted
for the multi-year tadpole and juvenile stages in MYL frogs
(Table S2).

@@

To identify important predictors of frog survival following
translocation, we developed linear and multilevel models in a Bayesian
framework. We performed all analyses with the rstanarm package
\citep{rstanarm2022} and R Statistical Software
\citep[v4.4.4,][]{rsoftware2022}, and used default, weakly informative
priors. Starting models included all relevant population-level effects
and, when data contained possible hierarchical structure, we included
group-level effects in the full model. We evaluated model fit using
leave-one-out (``LOO'') cross-validation
\citep{gelman2013, gabry2019, vehtari2016}, as implemented by the loo
package \citep{vehtari2022}.

We conducted two analyses. In the first, we addressed the more general
question of whether survival of frogs in the first translocation to a
site predicted the survival of frogs in subsequent translocations.
Group-level effects included site\_id or translocation\_id. The second
analysis focused specifically on identifying predictors of
post-translocation frog survival using a wide range of variables
describing characteristics of frogs, sites, and translocations
(predictor variables: Bd load (bd\_load), sex (sex\_male), frog size
(length), site elevation (elevation), winter severity in the year of
translocation (snow\_t), winter severity in the year following
translocation (snow\_t1), donor population (donor\_72996), day of year
on which a translocation was conducted (day), and translocation order
(order\_first)). In both analyses we used 1-year survival as a single
metric of survival that was available for all translocated populations
regardless of how many years of post-translocation data were available
(range = 3--16). Group-level effects included site\_id,
translocation\_id, or translocation\_id nested within site\_id. In both
analyses, the response variable, probability of 1-year frog survival,
was rounded to integer values to produce a binomial outcome, and all
models used a bernoulli family. For all models, we used default priors,
four chains, and 5000 iterations each for warmup and post-warmup. We
checked convergence of the MCMC algorithm using trace plots and Rhat.

@@

We analyzed skin swabs using standard Bd DNA extraction and qPCR methods
\citep[see SI Appendix for details]{boyle2004}. To assess the magnitude
of changes in Bd load on frogs following translocation, we compared Bd
loads measured before versus after translocation. Before-translocation
loads were quantified using skin swabs collected from all
to-be-translocated frogs at the donor site on the day before or the day
of the translocation. After-translocation Bd loads were based on all
swabs collected from translocated frogs at the recipient site in the
year of and the year following translocation. Individual frogs and their
associated Bd loads were included in the dataset only if frogs were
captured at the recipient site at least once during the year 0-1 period.
We used an analysis workflow similar to that described above. Briefly,
the response variable was before-translocation load and the predictor
variable was after-translocation load. We compared models with and
without group-level effects, and group-level effects included
pit\_tag\_id, pit\_tag\_id nested within site\_id, or pit\_tag\_id
nested within translocation\_id. All models used a negative binomial
family. Priors, chains, iterations, and checks were as described above.

Code to replicate all of the analyses described above is available from
the following GitHub repositories:
\url{https://github.com/SNARL1/translocation};
\url{https://github.com/SNARL1/cmr-analysis}.

@@

In this model we ignore density-dependent recruitment because we were
interested in the growth of the population from an initial
reintroduction and whether this growth was sufficient to prevent
extinction over 50 years following the introduction. We also did not
directly consider the dynamics of Bd in this model. We made this
decision because (i) translocated populations are infected with Bd at
high prevalence \citep{joseph2018}, and (ii) host density does not seem
to play a significant role in multi-year Bd infection dynamics in this
system \citep{wilber2022}. Thus, ignoring Bd infection dynamics and
instead assuming all host vital rates are in the presence of high Bd
prevalence significantly simplifies the model without much loss of
realism. Additional details are provided in the SI Appendix.

@@

\caption{\label{fig-allelefrequencies}Evidence for selection on
individual variants in recovering MYL frog populations at the landscape
scale. For each of the 9 naive and recovering MYL frog populations
(indicated by numbered points), adjacent pie charts show allele
frequencies for the 11 outlier SNPs from 7 distinct genes: (A)
LOC108802036, (B) TCF19, VARS, MAP1S, TYRP1, and SRFBP1, and (C) RIN3.
Charts are superimposed on a map of the Sierra Nevada study area, with
Yosemite, Kings Canyon, and Sequoia National Parks (from north to south)
shown in dark gray, and the range boundary for MYL frogs shown in light
gray. The inset map locates the study area and range boundary in
California.}

@@

{\centering \includegraphics[width=0.5\textwidth]{figures/translocation_survival_bysiteid.png}
